Predicting Solid State Material Platforms for Quantum Technologies

### 2021-12-13 paper.tex
Updates from Overleaf

Initially, we consider the materials that were classified as suitable with a probability of $85 \ \%$ or higher by all four ML models; BN, CdSe, BC$_2$N, InAs, CuI, and ZnCd$_3$Se$_4$. 
The compound BN (mp-$1639$) is already present in the training data as a suitable candidate, and therefore we believe the models recognize this as a suitable candidate with high probability. Furthermore, two compositions of CdSe (mp-$2691$ and mp-$1070$) have been predicted as suitable, possibly as a consequence of CdS being present as a suitable candidate in the training set. The element S resides in the same group as Se, and the data shows that the two compounds of CdSe share similar properties with calculated band gaps of $0.5$~eV and $0.55$~eV, respectively, in the MP database. 
Comparing to experiment, high quality nanocrystals of CdS have been synthesized and studied for use as quantum dots since the nanocrystals emit in the visible range with a direct band gap of $2.5$~eV \cite{CelebiSerdar2007SaCo, BanerjeeR2000Eots}. %Unfortunately, Cd is a toxic metal which could make large scale fabrication challenging.

In the case of BC$_2$N we find two compositions with the same chemical formula, the orthorhombic coordinated (mp-$629458$) and the chalcopyrite structured (mp-$1008523$) BC$_2$N, where the first takes on a polar space group while the latter does not. The band gaps are listed as $1.85$~eV and $1.65$~eV, respectively, in the MP database. BC$_2$N is known as heterodiamond and is a hybrid of diamond and BN. Both structures have, as expected, strong covalent character and have been studied for applications as nanostructures for electronic devices \cite{Gao2017}, hydrogen storage \cite{Cai2017} and superhard materials \cite{Li2017, Jiang2020}. It has similarly been predicted that the diamond-like structure of BC$_3$N can be a prominent spin qubit material host \cite{Wang2020SpinQB}. By creating a boron (B) vacancy one can obtain a nitrogen-vacancy center with similar properties to those found for the NV center in diamond. Whether this is also possible for BC$_2$N remains to be seen. We note that BC$_3$N is not present in the MP database and therefore not included in our dataset.

The compounds InAs (mp-$20305$), CuI (mp-$22895$ and mp-$569346$) and ZnCd$_3$Se$_4$ (mp-$1078597$) exhibit some comparable properties, and have band gaps of $0.30$, $1.18$ and $1.73$~eV, respectively. Single self-assembled InAs quantum dots have already been demonstrated \cite{Liu2018}, and is therefore an exciting possible material to use in quantum technology. To the best of our knowledge, ZnCd$_3$Se$_4$ has yet to be synthesized and contains the toxic element Cd, which could prove an obstacle for large scale utilization. CuI, however, has recently been synthesized and was shown to exhibit remarkable optoelectronic properties \cite{Ahn2020}. Interestingly, the material exhibits a large ionic character with more similarities to other oxides in the data set than, e.g., the more covalent materials of Si, SiC and diamond.  
It seems that ionic character alone is not an obstacle, but it is unknown at this time whether CuI has higher potential for hosting single-photon emitters, spin-manipulation or QT operation via nanostructuring. 
Note that NaCl was not predicted as suitable for QT in the intuitive approach.

Lowering the probability requirement from $85 \ \%$ to $75 \ \%$ for all models results in $66$ materials. 
The full list of these $66$ is contained in the Supplementary Material at \cite{supplementary}. 
The list of predicted suitable candidates now includes ternary compounds of the formula ABC$_2$. The elements Ga, Cd and Zn can occupy the A-site, Cu, Sn, Ag and Ge take the B-site, while S, Te, P or As may reside at the C site. Most of the formed compounds include toxic elements with one exception. This exception is ZnGeP$_2$ (mp-$4524$), which takes the chalcopyrite-like structure, with an indirect band gap of $1.2$~eV \cite{Zhang2015} as reported in the MP database and an experimentally reported band gap of $1.99$~eV \cite{Xing1989}.
ZnGeP$_2$ crystallizes in a non-polar space group, possesses no magnetic moment, has strong covalent bonds, and has been reported as an excellent mid-IR transparent crystal material that is suitable for nonlinear optical applications \cite{Zhang2015}. Importantly, it is possible to integrate sources of photon quantum states based on nonlinear optics with ZnGeP$_2$ \cite{Caspani2017}. 
We therefore identify ZnGeP$_2$ as an eligible candidate, but it remains unknown  
whether the candidate can facilitate the isolated deep energy levels often associated with defects showing quantum compatible properties. 
%if the candidate can provide isolation and shelter to experimentally facilitate a deep defect with quantum effects.

Among the $66$ materials that all models in the intuitive approach agreed on with more than $75 \ \%$ probability, we emphasize the presence of Ge, GeC, BP and InP. Ge in the cubic structure (mp-$1198022$) shares many properties with Si and C in addition to the periodic column number. 
%In fact, the first transistors were made in germanium due to their appealing electrical properties, but silicon took over as the material of choice for microelectronics due to the outstanding quality of silicon dioxide. 
%, which allowed the fabrication and integration of increasingly smaller transistors \cite{Scappucci2020, Pillarisetty2011}. 
Ge has the highest hole mobility of semiconductors at room temperature, and is therefore considered a key material for the process of extending the chip performance in classical computers beyond the limits imposed by miniaturization  \cite{Scappucci2020}. Furthermore, like SiC, Ge can take advantage of the mature large-scale fabrication of silicon due to their similar properties. 
Similar considerations could be made for the case of GeC. 
Data from the Materials Project database suggests that the cubic compound GeC (mp-$1002164$) is a covalent material having a band gap of $1.8$~eV. 
Interestingly, as SiC is widely known as highly suitable host material for QT compatible defects, we encourage further research on GeC due to its comparable properties with SiC.

The compound BP (mp-$1479$ and mp-$1008559$) is present in our predictions in both the cubic and hexagonal structures. The indirect band gaps of the different structures are calculated at $1.46$~eV and $1.1$~eV, respectively, as reported in the MP database. They are both nonmagnetic and share several properties with the entries mentioned above; including ... The compound is non-toxic and has been synthesized with a potential for large-scale production \cite{MukhanovVladimirA2016Umso}.

Lastly, we will mention the prediction of InP (mp-$966800$) as a suitable candidate. The compound inhabits the hexagonal structure with corner-sharing InP$_4$ tetrahedra. InP is reported in the Materials Project to have a direct band gap of $0.51$~eV, and is considered as one of the most promising candidates 
to compete with 
Cd- or Pb- based QDs for, e.g., display and lighting applications \cite{Zhang2020a, Won2019}.

It is worth to mention that all models agree on several oxides being potential candidates when the probability cut-off is set low enough (but still above $50 \ \%$). However, we find that almost all the oxide compounds fall between the decision boundary defining suitable and unsuitable candidates, and none are present in the list defining the $66$ suitable candidates in the Supplementary Material. Due to the labeling of ZnO as a suitable candidate, we believe that the boundary was shifted sufficiently to admit several oxides as suitable candidates.
Removing ZnO from the training set for the intuitive approach might result in fewer or even no oxide compounds being predicted above the suitability limit.

The work of \citeauthor{Ferrenti2020} \cite{Ferrenti2020} suggests a list of $541$ viable hosts after the data mining procedure.  
Of these, we find only a single material present in our list of $66$ candidates (see the Supplementary Material at \cite{supplementary}). This material is the nontoxic compound MgSe (mp-$10760$) which crystallizes in the rock-salt structure. Its band gap is $1.98$~eV in the MP database with the experimental value being $5.6$~eV  \cite{SaumGeorge1959}. MgSe is notable for its available spin-zero isotopes in accordance with the criteria set by the of Ref.~\cite{Ferrenti2020}. We believe that these criteria favor defects acting as spin centers with qubit potential and MgSe is thus identified as a promising host material in this respect.  
%\oliver{Maybe also see if other compounds that were present in training set are present in their 541 hosts. }

%Supplementary information: List displaying the 66 predicted candidates that all models in the intuitive approach agreed on with more than 75\% probability. %appendix/listof66materials

@@

Out of the three approaches, we find that the augmented Ferrenti approach is the least restricted approach and admits the most entries. The Ferrenti approach also admits a large number of entries, but is otherwise not very different from the augmented version. 
Importantly, by applying the four ML methods to the data extracted using the two approaches, we are unable to reproduce the criteria the approaches are based on, such as band gap restriction or polar space group. Of course, the list of materials proposed by the Ferrenti and augmented Ferrenti approach is extensive, whereas the intuitive approach predicts fewer suitable candidates where manually verification is achievable.  
Manual verification through a literature survey will often not be available, however, and perfecting automatic data mining and analysis is therefore an important goal of material informatics \cite{rickman2019}.

It is important to note that the intuitive approach also proposed candidates with band gap energies lower than $0.5$~eV, but less compared to that of the other two approaches. Thus, all three approaches predicted suitable candidates with band gaps lower than the lowest in the training sets. 
%We believe there are three reasons for this.  
It is known that the GGA functional Materials Project applies is known to underestimate the band gap of many semiconductors. 
%underestimating the band gap severely, and therefore there is a chance that the models consider the band gap as noise and not useful information. 
It is therefore possible that the band gap is not considered an important parameter by the ML methods. 
%Understanding the motivation behind the classification of the models will be an important step towards understanding ... 
Indeed, we did not find that the band gap was prominent in the principal components, although the band gap could be correlated with other more important features. 
Thus, one can argue that the ML methods find other patterns that represent a better distinction between suitable and unsuitable candidates in the training sets, resulting in the band gap being redundant.

We find that $119$ of the $214$ candidates predicted as suitable by the intuitive approach (when the probability cut-off is set to [X]) were also selected by all ML methods in the augmented Ferrenti approach. 
%Of the $214$ suitable candidates predicted by all models in the intuitive approach, we find $119$ of them also predicted as suitable by all models in the augmented Ferrenti approach. 
Similarly, $78$ of the $214$ are also predicted as suitable by all models in the Ferrenti approach. All approaches and their corresponding methods agree on a $47$ potential candidates, where eight are elementary (unary), $29$ binary, and $10$ tertiary.
\textbf{The list of the $47$ materials agreed on by all ML methods and all three approaches as suitable candidates with a probability $> X \ \%$ can be found in the Supplementary Material at \cite{supplementary}.}

\section*{Discussion} 
% Shorter section, includes summary, concluding remarks and is quite heavy on outlook.

%\oliver{The following paragraphs are "disclaimers" where we list possible pitfalls and weaknesses of this experiment. Could also be incorporated in the discussion instead?} MB I moved it up to data mining

%SHORT SUMMARY FIRST HERE.

Motivated by our findings, we believe that further computational, experimental, or theoretical verification after a prediction of a possible promising material remains an important step in this work. This step is part of the workflow for novel materials discovery, which is visualized in Fig.~\ref{fig:ht-workflow}.  
Nevertheless, we have provided an exploratory analysis for the discovery of novel materials to be used in QT. Considering the number of materials predicted as suitable candidates by our machine learning methods and approaches, we hope it encourages further studies and identification of possible new material hosts.

We find that including toxic materials etc in the training set is vital for reliable predictions. They can then later be excluded when planning experiments, but the practical considerations should be saved for last and not included in the initial criteria.

\section*{Methods}
%Should be rather short. Details can be included in a Supplementary Material.

\subsection*{Databases} %Sebastian and Oliver 
The Materials Project \cite{Jain2013, Jain2018} is an open-source project based on the Vienna Ab Initio Package (VASP) \cite{Kresse1996}. The Perdew-Burke-Ernzerhof \cite{Perdew1996} (PBE) functional is used to calculate band structures, while for transition metals, a $+U$ correction is applied to correct for correlation effects \cite{Wang2006}. The project is known as the initiator of materials genomics and offers a variety of calculated properties for over one hundred thousand inorganic crystalline materials, with frequent updates and extensions.

The Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015} contains thermodynamic and structural properties of more than 600.000 materials. The calculations are performed with the VASP software whereas the electron exchange and correlation are described with the PBE functional. The $+U$ extension is included for several calculations considering specific elements \cite{Stevanovic2012}.

JARVIS-DFT \cite{Choudhary2020} is an open database based on the VASP software and consists of roughly 40.000 3D materials using the vdW-DF-OptB88 functional \cite{Thonhauser2007, Klimes2011}. Structures included in the database are originally taken from the Materials Project, and then re-optimized using the OPT-functional. Finally, the combination of the OPT and modified Becke-Johnson (mBJ) functionals is used to obtain a representative band gap of each structure \cite{Choudhary2018a}.

@@

AFLOW-ML \cite{Isayev2017} is an API that uses machine learning to predict thermo-mechanical and electronic properties based on the chemical composition and atomic structure alone, which they denote as \textit{fragment descriptors}. Initially, the API decides whether a given material is a metal or an insulator, where the latter is confirmed with an additional regression method to predict the band gap width. The accuracy is validated by a five-fold cross-validation process for each ML method, where they report a $93 \ \%$ prediction success of their initial binary classification method.

Matminer \cite{Ward2018} is an open-source toolkit for material analysis written in Python. Matminer provides modules to extract information from a wide variety of databases. Additionally, they provide the tools to construct possibly thousands of features from calculations based on a material's composition, structure and electronic properties from DFT calculations, and have frameworks for visualization and automatic machine learning.

\subsection*{Material Informatics}  
Screening and workflow and approaches. Technicalities of featurization if necessary.

@@

In this work we have applied four well-known and tested machine learning methods for classification problems, these are (see for example \cite{Hastie2009,Mehta2019} for discussions and applications):
\begin{enumerate}
    \item Logistic regression,
    \item Decision trees,
    \item Random forests,
    \item Gradient boosting.
\end{enumerate}
Logistic regression \cite{Hastie2009} is a simple and frequently used method for binary and multi-category classification problems.  In addition to logistic regression, we have also applied and tested the predictions made by decision trees and ensemble methods like random forests and gradient boosting, the latter through the application of the computationally efficient  XGBoost library \cite{xgboost2016}. Gradient boosting and random forests use decision trees as weak learners and improve their predictability. For random forests this is implemented through a collection of randomized decision trees where a  subset of the features in the data sets are selected randomly when building a decision tree. Boosting methods like gradient boosting use decision trees as  weak learners and improve upon these by an iterative process that involves the estimation of the gradients of the cost/loss function \cite{Hastie2009}. Pure decision trees can easily lead to overfitting of the data under study, leading to a machine learning method that exhibits a high variance. Ensemble methods like random forests and gradient boosting on the other hand tend to soften the overfitting problem, resulting in both a small bias and a reduced variance of the employed method, see for example Refs.~\cite{Hastie2009,Mehta2019} for an in-depth discussion of the bias-variance trade-off in machine learning. Gradient boosting implemented through the  XGBoost library \cite{xgboost2016}, is widely used by data scientists to achieve state-of-the-art results on many machine learning challenges.

\section*{Data availability} 
%We can put something else, standard text 
%The data that support the plots within this paper and other findings of this study are available from the corresponding authors upon reasonable request. [Marianne: I assume you want to add the Github link here?]

The data sets that support the findings of this study are available from the corresponding author, and the data sets used in our  machine learning studies are available online at [Github link].

\section*{Code availability} 
%The codes developed in this study are available from the authors upon reasonable request. [Marianne: I assume you want to add the Github link here?] 
The codes employed to develop the machine learning results are available online at https://github.com/ohebbi/predicting-solid-state-qubit-material-hosts (DOI: 10.5281/zenodo.4633959 - Update with new release later). %This DOI refers to an release of a Github Repository.

\bibliography{apssamp}% Produces the bibliography via BibTeX.

>>>>

Initially, we consider the materials that were classified as suitable with a probability of $85 \ \%$ or higher by all four ML models; BN, CdSe, BC$_2$N, InAs, CuI, and ZnCd$_3$Se$_4$. 
The compound BN (mp-$1639$) is already present in the training data as a suitable candidate, and therefore we believe the models recognize this as a suitable candidate with high probability. Furthermore, two compositions of CdSe (mp-$2691$ and mp-$1070$) have been predicted as suitable, possibly as a consequence of CdS being present as a suitable candidate in the training set. The element S resides in the same group as Se, and the data shows that the two compounds of CdSe share similar properties with calculated band gaps of $0.5$~eV and $0.6$~eV, respectively, in the MP database. 
Comparing to experiment, high quality nanocrystals of CdS have been synthesized and studied for use as quantum dots since the nanocrystals emit in the visible range with a direct band gap of $2.5$~eV \cite{CelebiSerdar2007SaCo, BanerjeeR2000Eots}.

In the case of BC$_2$N we find two compositions with the same chemical formula, the orthorhombic coordinated (mp-$629458$) and the chalcopyrite structured (mp-$1008523$) BC$_2$N, where the first takes on a polar space group while the latter does not. The band gaps are listed as $1.9$~eV and $1.7$~eV, respectively, in the MP database. BC$_2$N is known as heterodiamond and is a hybrid of diamond and BN. Both structures have, as expected, strong covalent character and have been studied for applications as nanostructures for electronic devices \cite{Gao2017}, hydrogen storage \cite{Cai2017} and superhard materials \cite{Li2017, Jiang2020}. It has similarly been predicted that the diamond-like structure of BC$_3$N can be a prominent spin qubit material host \cite{Wang2020SpinQB}. By creating a boron (B) vacancy one can obtain a nitrogen-vacancy center with similar properties to those found for the NV center in diamond. Whether this is also possible for BC$_2$N remains to be seen. We note that BC$_3$N is not present in the MP database and therefore not included in our dataset.

The compounds InAs (mp-$20305$), CuI (mp-$22895$ and mp-$569346$) and ZnCd$_3$Se$_4$ (mp-$1078597$) have band gaps of $0.3$, $1.2$ and $1.7$~eV, respectively, and exhibit comparable properties. Self-assembled InAs quantum dots, for example, are an exciting possible material to use in quantum technology \cite{Liu2018}. To the best of our knowledge, ZnCd$_3$Se$_4$ has yet to be synthesized and contains the toxic element Cd, which could prove an obstacle for large scale utilization. CuI, however, has recently been synthesized as single-crystal epitaxial films and was shown to exhibit remarkable optoelectronic properties \cite{Ahn2020}. Interestingly, the material exhibits a large ionic character with more similarities to other oxides in the data set than, e.g., the more covalent materials of Si, SiC and diamond.  
It seems that ionic character alone is not an obstacle, but it is unknown at this time whether, if CuI in fact does have potential for hosting single-photon emitters and spin centers, these properties would result from deep level defects or nanostructuring. 
Note that NaCl was not predicted as suitable for QT in the intuitive approach.

Lowering the probability requirement from $85 \ \%$ to $75 \ \%$ for all models results in $66$ materials. 
The full list of these $66$ is contained in the Supplementary Material at \cite{supplementary}. 
The list of predicted suitable candidates now includes ternary compounds of the formula ABC$_2$. The elements Ga, Cd and Zn can occupy the A-site, Cu, Sn, Ag and Ge take the B-site, while S, Te, P or As may reside at the C site. Most of the predicted compounds include toxic elements with one exception: ZnGeP$_2$ (mp-$4524$) in a chalcopyrite-like structure with an indirect band gap of $1.2$~eV \cite{Zhang2015} as reported in the MP database. In comparison, the experimentally reported band gap is larger at $2.0$~eV \cite{Xing1989}.
ZnGeP$_2$ crystallizes in a non-polar space group, possesses no magnetic moment, exhibits strongly covalent bonds, and has been reported as an excellent mid-IR transparent crystal material that is suitable for nonlinear optical applications \cite{Zhang2015}. Importantly, it is possible to integrate sources of photon quantum states based on nonlinear optics with ZnGeP$_2$ \cite{Caspani2017}. 
We therefore identify ZnGeP$_2$ as an eligible candidate, but it remains to be seen whether the candidate can facilitate the isolated deep energy levels often associated with defects showing quantum compatible properties.

Among the $66$ materials that all models in the intuitive approach agreed on with more than $75 \ \%$ probability, we emphasize the presence of Ge, GeC, BP and InP. Ge in the cubic structure (mp-$1198022$) shares many properties with Si and C in addition to the periodic column number. 
Ge has the highest hole mobility of all semiconductors at room temperature, and is therefore considered a key material for the process of extending the chip performance in classical computers beyond the limits imposed by miniaturization \cite{Scappucci2020}. Furthermore, like SiC, device design based on Ge can take advantage of the mature large-scale fabrication of silicon due to the material's comparable properties. 
Similar considerations could be made for the case of GeC. 
Data from the Materials Project database suggests that the cubic compound GeC (mp-$1002164$) is a covalently bonded semiconductor having a band gap of $1.8$~eV. 
Consequently, as SiC is widely known as highly suitable host material for QT compatible defects, we encourage further research on GeC due to its similarities to SiC.

The compound BP (mp-$1479$ and mp-$1008559$) is present in our predictions in both the cubic and hexagonal structures. The indirect band gaps of the different structures are calculated at $1.5$~eV and $1.1$~eV, respectively, as reported in the MP database. Both configurations of PB are nonmagnetic and share several properties with the entries mentioned above; including ... The compound is non-toxic and has been synthesized with a potential for large-scale production \cite{MukhanovVladimirA2016Umso}.

Lastly, we will mention the prediction of InP (mp-$966800$) as a suitable candidate. The compound inhabits the hexagonal structure with corner-sharing InP$_4$ tetrahedra. InP is reported in the Materials Project to have a direct band gap of $0.5$~eV and is considered as one of the most promising candidates to compete with 
Cd- or Pb- based QDs for, e.g., display and lighting applications \cite{Zhang2020a, Won2019}. 
We note the possibility of using InP-based quantum dots for QT application.

It is worth to mention that all models agree on several oxides being potential candidates when the probability cut-off is set low enough (but still above $50 \ \%$). However, we find that almost all the oxide compounds fall between the decision boundary defining suitable and unsuitable candidates, and none are present in the list defining the $66$ suitable candidates in the Supplementary Material. Due to the labeling of ZnO as a suitable candidate, we believe that the boundary was shifted sufficiently to admit several oxides as suitable candidates. 
Removing ZnO from the training set for the intuitive approach might result in fewer or even no oxide compounds being predicted above the suitability limit. 
The work of \citeauthor{Ferrenti2020} \cite{Ferrenti2020} suggests a list of $541$ viable hosts after the data mining procedure.  
Among these these, we find only a single material present in our list of $66$ candidates predicted by the four ML models after training on the dataset resulting from the intuitive approach (see the Supplementary Material at \cite{supplementary}). This material is the nontoxic compound MgSe (mp-$10760$) which crystallizes in the rock-salt structure. Its band gap is $2.0$~eV in the MP database with the experimental value being reported as much larger at $5.6$~eV  \cite{SaumGeorge1959}. MgSe is notable for its available spin-zero isotopes in accordance with the criteria set by the of Ref.~\cite{Ferrenti2020}. We believe that these criteria may favor defects acting as spin centers with qubit potential and MgSe is thus identified as a promising host material in this respect.  
%\oliver{Maybe also see if other compounds that were present in training set are present in their 541 hosts. }

@@

Out of the three approaches, we find that the augmented Ferrenti approach is the least restricted approach and admits the most entries. The Ferrenti approach also admits a large number of materials into the training set, but the resulting data is otherwise not very different from the augmented version. 
Importantly, by applying the four ML methods to the data extracted using the two approaches, we are unable to reproduce the criteria the approaches are based on, such as band gap restriction or polar space group. Of course, the list of materials proposed by the Ferrenti and augmented Ferrenti approach is extensive, whereas the intuitive approach predicts fewer suitable candidates where manually verification is achievable.  
Manual verification through a literature survey will often not be possible, however, and perfecting automatic data mining and analysis is therefore an important goal of material informatics \cite{rickman2019}.

It is important to note that the intuitive approach also proposed candidates with band gap energies lower than $0.5$~eV, but to a lesser extent as compared to the other two approaches. Thus, all three approaches predicted suitable candidates with band gaps lower than the lowest in the training sets. 
Firstly, as mentioned above, it is known that the PBE functional that the Materials Project applies is known to underestimate the band gap of many semiconductors \cite{Freysoldt2014}. 
It is therefore possible that the band gap is not considered an important parameter by the ML methods. 
Indeed, we did not find that the band gap was prominent in the principal components, although the band gap could be correlated with other more important features. 
Thus, one can argue that the ML methods are finding  other patterns that represent a better distinction between suitable and unsuitable candidates in the training sets, resulting in the band gap being redundant.
\textbf{In this context, we also note that the patterns recognized by the ML algorithms are challenging to discern. Thus, while we intend for the methods to predict materials for QT, we cannot be certain that this is happening. 
However, the large degree admittance of materials present in the training set of all three approaches (Ferrenti, augmented Ferrenti and intuitive)  through to the predictions indicates that at least some of the intended parameters are being recognized as important.  
Among the three approaches to data mining we find that the intuitive one conforms to expectation most closely. }

\textbf{I am missing a discussion, here or below, on the different ML methods and their different results. Which of the four algorithms is best or makes the most sense? Please compare either here, or for each of the approaches in the Predictions sections.}

We find that $119$ of the $214$ candidates predicted as suitable by the intuitive approach (when the probability cut-off is set to $50 \ \%$) were also selected by all ML methods in the augmented Ferrenti approach. 
Similarly, $78$ of the $214$ are also predicted as suitable by all models in the Ferrenti approach. All approaches and their corresponding methods agree on a total of $47$ potential candidates, where eight are elementary (unary), $29$ binary, and $10$ tertiary.
\textbf{The list of the $47$ materials agreed on by all ML methods and all three approaches as suitable candidates with a probability $> 50 \ \%$ can be found in the Supplementary Material at \cite{supplementary}.}

\section*{Discussion} 
We have developed data extraction tools for numerous material science databases, and constructed over $4800$ physics-informed features for a data set consisting of more than $25000$ materials. Furthermore, three approaches to data mining for QT applications were developed; the Ferrenti, augmented Ferrenti and augmented approaches. 
Three distinct training sets were defined for four different machine learning algorithms (logistic regression, decision tree, random forest and gradient boost).

We find that the formulation of the training sets for the Ferrenti and augmented Ferrenti approaches was overly broad with a lack of clear distinction between the materials labeled suitable and unsuitable for QT in the two training sets. The stringent restrictions set for the intuitive approach on the other hand proved more suitable basing on the much clearer distinction between suitable and unsuitable candidates.

Notably, our results show that including impractical materials (e.g., toxic, radioactive and without spin-zero isotopes) in the training set is vital for defining a suitable training set. 
\textbf{This manifests in the ML predictions as ...} 
Such materials can be eliminated based on practical considerations at a later stage but should be included initially in order to provide the ML methods with ideal conditions to divine physics based trends.

In short, the findings presented herein firmly establish that material informatics via data mining to machine learning predictions is a both viable and important route to new discoveries in important fields. Herein, we focus on predicting new candidate materials to host single-photon emitters and spin centers for quantum technology applications, but the framework we have developed is suitable for other fields as well. 
Two main routes can be followed to exploit the results. One aspect is the pursuit of experimental verification of quantum technology compatible effect in the materials predicted as suitable by the four ML methods.  
Another, perhaps even more important, is to use the features and trends identified during the data mining and prediction process to understand the distinct material characteristics that enable quantum effects to manifest. Thus can new discoveries in the field of quantum technologies be made.

\section*{Methods}

\subsection*{Databases} 
The Materials Project \cite{Jain2013, Jain2018} is an open-source project based on the Vienna Ab Initio Package (VASP) \cite{Kresse1996}. The Perdew-Burke-Ernzerhof \cite{Perdew1996} (PBE) functional is used to calculate band structures, while for transition metals, a $+U$ correction is applied to correct for correlation effects \cite{Wang2006}. The project is known as the initiator of materials genomics and offers a variety of calculated properties for over one hundred thousand inorganic crystalline materials, with frequent updates and extensions.

The Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015} contains thermodynamic and structural properties of more than $600.000$  materials. The calculations are performed with the VASP software whereas the electron exchange and correlation are described with the PBE functional. The $+U$ extension is included for several calculations considering specific elements \cite{Stevanovic2012}.

JARVIS-DFT \cite{Choudhary2020} is an open database based on the VASP software and consists of roughly $40.000$ 3D materials using the vdW-DF-OptB88 (OPT) functional \cite{Thonhauser2007, Klimes2011}. Structures included in the database are originally taken from the Materials Project, and then re-optimized using the OPT functional. Finally, the combination of the OPT and modified Becke-Johnson (mBJ) functionals is used to obtain a representative band gap for each structure \cite{Choudhary2018a}.

@@

AFLOW-ML \cite{Isayev2017} is an API that uses machine learning to predict thermo-mechanical and electronic properties based on the chemical composition and atomic structure alone, which are denoted as \textit{fragment descriptors}. Initially, the API decides whether a given material is a metal or an insulator, where the latter is confirmed with an additional regression method to predict the band gap width. The accuracy is validated by a five-fold cross-validation process for each ML method, where they report a $93 \ \%$ prediction success of their initial binary classification method.

\subsection*{Material Informatics}  
Matminer \cite{Ward2018} is an open source toolkit for material analysis written in Python. Matminer provides modules to extract information from a wide variety of databases. Additionally, they provide the tools to construct possibly thousands of features from calculations based on a material's composition, structure and electronic properties from DFT calculations, and have frameworks for visualization and automatic machine learning. 
The features selected for featurization herein are summarized in the Supplementary Material at \cite{supplementary}.

@@

In this work we have applied four well-known and tested machine learning methods for classification problems, these are (see for example \cite{Hastie2009,Mehta2019} for discussions and applications):
\begin{enumerate}
    \item Logistic regression,
    \item Decision trees,
    \item Random forests,
    \item Gradient boosting.
\end{enumerate}
Logistic regression \cite{Hastie2009} is a simple and frequently used method for binary and multi-category classification problems. In addition to logistic regression, we have also applied and tested the predictions made by decision trees and ensemble methods like random forests and gradient boosting, the latter through the application of the computationally efficient  XGBoost library \cite{xgboost2016}. Gradient boosting and random forests use decision trees as weak learners and improve their predictability. For random forests this is implemented through a collection of randomized decision trees where a  subset of the features in the data sets are selected randomly when building a decision tree. Boosting methods like gradient boosting use decision trees as  weak learners and improve upon these by an iterative process that involves the estimation of the gradients of the cost/loss function \cite{Hastie2009}. Pure decision trees can easily lead to overfitting of the data under study, leading to a machine learning method that exhibits a high variance. Ensemble methods like random forests and gradient boosting on the other hand tend to soften the overfitting problem, resulting in both a small bias and a reduced variance of the employed method, see for example Refs.~\cite{Hastie2009,Mehta2019} for an in-depth discussion of the bias-variance trade-off in machine learning. Gradient boosting implemented through the  XGBoost library \cite{xgboost2016} is widely used by data scientists to achieve state-of-the-art results on many machine learning challenges.

\section*{Data availability} 
The data sets that support the findings of this study are available online at DOI: 10.5281/zenodo.4633959.

\section*{Code availability} 
The codes employed to develop the machine learning results are available online at 
%https://github.com/ohebbi/predicting-solid-state-qubit-material-hosts 
DOI: 10.5281/zenodo.4633959. 
%Update with new release later).

\bibliography{apssamp}
